---
title: "Project"
author: "Harshali Gaikwad"
date: "2024-11-09"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)
```


```{r}
library(corrplot)
library(lmtest)
library(car)
library(MASS)
library(lmtest)


medical_insurance_data <- read.csv("insurance.csv", header = TRUE)
head(medical_insurance_data)
```
Lets perform EDA on the data
                                                                                                                   
```{r}
hist(medical_insurance_data$age, 
     main = "Distribution of Age",
     xlab = "Age",
     ylab = "Count",
     col = "skyblue",
     border = "white")

# Calculate the counts for each sex category
sex_counts <- table(medical_insurance_data$sex)

# Define colors for each slice of the pie
colors <- c("#FF9999", "#66B2FF")

# Calculate percentages for each category
percentages <- round(100 * sex_counts / sum(sex_counts), 1)

# Create labels that include both the name and percentage
labels <- paste0(names(sex_counts), ": ", percentages, "%")

# Plot the pie chart
pie(sex_counts,
    main = "Sex Distribution",
    col = colors,
    labels = labels,
    cex = 1.2,
    radius = 0.8)

# Add a legend for better interpretation
legend("topright", legend = paste0(names(sex_counts), " (", sex_counts, ")"),
       fill = colors, cex = 0.8, bty = "n")

hist(medical_insurance_data$bmi, 
     main = "Distribution of BMI",
     xlab = "BMI",
     ylab = "Count",
     col = "darkgreen",
     border = "white")
```



```{r}
medical_insurance_data$sex <- ifelse(medical_insurance_data$sex == "male", 1, 0)
medical_insurance_data$smoker <- ifelse(medical_insurance_data$smoker == "yes", 1, 0)
medical_insurance_data$region <- as.numeric(factor(medical_insurance_data$region))

head(medical_insurance_data)

correlation_matrix <- cor(medical_insurance_data)
corrplot(correlation_matrix, method = "color", type = "full", 
         col = colorRampPalette(c("blue", "white", "red"))(200), 
         title = "Correlation Heatmap", 
         mar = c(0,0,1,0))

pairs(medical_insurance_data)


```

```{r}
model1 <- lm(charges ~ .,data = medical_insurance_data)
summary(model1)
```

From the summary of the model we find that the attribute 'sex' is not significant so we will try a model without this attribute.

```{r}
model2 <- lm(charges ~ age + bmi + children + smoker + region, data = medical_insurance_data)
summary(model2)
```

#```{r}
model3 <- lm(charges ~ age + smoker + bmi + bmi * smoker + children + region, data = medical_insurance_data)
summary(model3)
#```

```{r}
model3 <- lm(charges ~ age + smoker + bmi + sex + sex * smoker + children + region, data = medical_insurance_data)
summary(model3)
```

We get the best model when we included the interaction term. Lets check the autocorrelation test
```{r}
dwtest <- dwtest(model3)
dwtest

residuals <- residuals(model3)

acf(residuals, main="ACF of Residuals")
```


The diagram shows that there is no autocorrelation.Moreover, the dwtest tell us that there is no significant autocorrelation between the residuals of the model.

Since there is no autocorrelation, lets check for multicollinearity using the VIF test

```{r}
vif(model3)
```

Looking at the vif we find that there is no multicollinearity. \
Now we check for the homoscedasticity by plotting the model.

```{r}
par(mfrow=c(2,2))
plot(model3)
```


```{r}
bc <- boxcox(model3)
lambda <- bc$x[which.max(bc$y)] 

if (abs(lambda) < 1e-4) {  # If lambda is close to 0
    transformed_charges <- log(medical_insurance_data$charges)
} else {
    transformed_charges <- (medical_insurance_data$charges^lambda - 1) / lambda
}

model3_transformed <- lm(transformed_charges ~ age + smoker + bmi + sex + sex * smoker + children + region, data = medical_insurance_data)

summary(model3_transformed)

plot(model3_transformed)
```

As we can see in the standardized residual vs fitted plot, it is better than the previous models plot. It is some what randomly distributed.